Applying State-free Reversible VAMPNets and Markov State Models to Learn Dynamics of DNA Oligonucleotides

### 2020-12-07 manuscript.tex
Updates from Overleaf
move f4 to supplement, added more refs, fixed fig size

% need to discuss tmelt shift
We measured the slow dissociation response by fitting the distribution of times at which the core base pairs separate beyond a 2 nm cutoff. We found the inverse of these relaxation times -- the effective dissociation rate -- to increase exponentially with temperature, which is expected given the large enthalpic barrier of dissociation. We noted an average temperature shift of about 4C between experiment and simulation exponentials. While 3spn2 captures melting temperature relatively well, some deviation is expected given varied ion conditions and other coarse-grained effects. We used to this temperature shift to inform our equilibrium simulations in the next section. We also noted an acceleration of about one order of magnitude compared to experiment, although this factor is sensitive to the exact definition of melting temperature. After accounting for the melting temperature shift and acceleration factor, we saw strong agreement between the experimental data and simulated relaxation fits. At lower temperatures, T-jump slow responses likely contain a mix of dissociation and hybridization dynamics, therefore rates do not drop off to the same extent as our analysis which considers dissociation alone. GC-core showed the largest deviation from experiment, indicating that increased temperature triggers dissociation at a disproportionately higher rate in simulation compared to experiment.

@@

For A:T terminal sequences, the simulated fast response appears linear with temperature, indicating a barrierless and diffusion-driven process. In contrast, the GC-end responses are distinctly slower and increase exponentially with temperature, likely due to a greater thermodynamic barrier associated with G:C fraying. We observe similar trends in the experimental data, although comparisons at high temperatures were limited due to mixing between the fast and slow responses (figure \ref{fig:relaxation-comparison}). We found the optimal acceleration factor to be dependent on the choice of the cutoff parameter, but we can approximate that 3spn2 fraying dynamics are accelerated by about 100-200x relative to experiments. It is not surprising that we see different rates of acceleration for the dissociation and fraying processes given that coarse-grained effects can vary across different degrees of freedom. In particular, the simplified treatment of the fraying process may smoothen the free energy landscape and speed up dynamics relative to a more global processes like dissociation. Although one should rely on higher resolution models to study in depth mechanisms of fraying, our results indicate that terminal fraying is a reasonable assignment for the spectroscopic fast response in Sanstead et al. and that 3spn2 can capture some sequence-dependent fraying effects.
% need a better transition here back into SRV-MSMs?

@@

For GC-end we observe no significant flux between the 5S2 state and the structurally similar 5S1 state, indicating that the former acts more as a kinetic trap than a pathway to H. The 5S1 and 3S1 states still readily convert with H, however there is an order of magnitude lower probability of reaching these states from D when compared to AT-all. This indicates that inert tails inhibit out-of-register binding -- accounting in part for higher free energies discussed above -- but may not substantially disrupt out-of-register hybridization mechanisms along 5S1 $\rightarrow$ H and 3S1 $\rightarrow$ H once shifted states have formed.

@@

Taken together, our results indicate that AT-all hybridization and dissociation kinetics are substantially modulated by out-of-register pathways. A majority of hybridization events occur out-of-register, and, based on transition probabilities, 35\% of D $\rightarrow$ H pathways and 37\% of H $\rightarrow$ D pathways pass through some out-of-register state. The more shifted 3S2 and 5s2 states are more prone to dissociation and thus contribute to about 5\% of these pathways. While GC-end strands can access out-of-register states, only 11\% of native hybridization events pass through any out-of-register states. These observations show that internal displacement mechanisms reported in previous simulation studies are capable of disrupting non-native pairing and correcting base pair mismatches without fully dissociating strands \citep{Romano2013DNADependence, Markegard2015, Maciejczyk2014DNAModel}. Experimentally, out-of-register states are suspected to occur, but are difficult to capture due to short lifetimes and response similarity to other processes such as fraying. To explicitly minimize out-of-register base pairing, similar AT repeat motif sequences have been padded by GCG clamps during experimental analysis \cite{Wyer2014KineticsAT-tracts}. Recent all-atom results identified similar out-of-register states as "deep kinetic traps" along the hybridization pathway for repetitive dGCGCGC hexamers \citep{Xiao2019}. Contrary to our kinetic results, however, "slithering" mechanisms along out-of-register tracts were not observed to be a dominant pathway, especially when compared to the high rates of slithering exhibited by the homogenous dGGGGGG strand. It is unclear whether these differences were a consequence of varied simulation conditions or whether GC repeat motifs are less susceptible to out-of-register transitions compared to AT motifs. This is possible given that stronger hydrogen binding in GC motifs may prevent fluctuation-driven rearrangement, however further computational and experimental studies are required to verify these differences.

\subsection{Stretched T-jump IR responses correspond to shifting sequences}

@@

% discuss computational comparisons needs work
Lattice model studies have shown that frayed intermediates make substantial contributions to the GC-core conformational ensemble \citep{Araque2016LatticeCooperativity, Phys2019}. Araque et al. defines a similar 8-mer sequence (dATGCGCAT) as non-two-state, where a stable, symmetrically A:T frayed state is a crucial part of the duplex transition path \citep{Araque2016LatticeCooperativity}. <Add Brennan example> When examining all four sequences using T-jump IR and 2D IR spectroscopy, Sanstead et al. found that the GC-core had the highest deviation from two-state behavior during dissociation (when excluding out-of-register contributions) \citep{Sanstead2016}. As T-jump IR data only showed a loss of A:T contacts during the fast response, this intermediate state was defined by a high degree of fraying about the central core. While 1-2 base pair fraying was commonly observed for GC-mix and AT-all as well, lattice model predictions showed that GC-core had substantially more frayed base pairs \citep{Phys2019}. Variable T-jump measurements and Smoluchowski simulations on model 1D free energy landscapes showed that AT termini fraying was an effectively barrierless process characterized by rapid inter-conversion between all accessible frayed states \citep{Sanstead2018DirectDehybridization}. We see the same rapid fraying in simulation data -- which is too fast to be attributed to a converged SRV mode -- however our SRV-MSM results show this inter-conversion first relies on the slower formation of the the A:T bond nearest to the GC center.  Although this process occurs much slower than single A:T base bonding and breaking, it may be difficult to experimentally discern from the overall hybridization process which contains both G:C and A:T character and occurs on a similar timescale.

@@

\subsection{Macrostate distributions inform state stability and kinetics}

% add new diffusion map section introduction
Although our SRV-MSMs provide useful insight into slow processes, we were also interested in the ensemble of configurations within PCCA+ macrostates. These configurations interchange significantly faster than the SRV and MSM lag times, but we can use a diffusion map embedding to glean a structural understanding of the macrostate population. Diffusion maps generate a low dimensional embedding of the data into high variance structural modes and are well-suited to find subtle differences in temporally disconnected data \citep{Coifman2006DiffusionMaps, Ferguson2010SystematicMaps}. We were particularly interested in using this method to explore the 5s1 and F4 macrostates as these represent the most populated and kinetically relevant macrostates for AT-all, GC-end, and GC-core.

@@

@@

The first diffusion mode clearly delineates between the GC-end and AT-all shifted conformations and correlates highly with the average distance between the 3' end and its shifted complementary pair. As expected, this shows that mismatched C:T pairs are never bound whereas the AT-all pairs are mostly bound with occasional fraying indicated by small AT-all overlap in the GC-end region. The second diffusion mode, which correlates highly with the average distance between 3' and 5' ends, highlights structural variances within the two sequences distributions. We notice that about 40\% of GC-end configurations fall below $\Psi_3 = -0.01$  compared to only 10\% of AT-all configuration. We find that these GC-end populations share stabilizing termini contacts despite all A:T contacts being shifting out-of-register. These "shifted-loop" bonds appear uniquely stable for GC-end conformations,  and suggest an intermediate states through which hybridization is facilitated. This may account for the GC-end 5S1 $\rightarrow$ H transition being more favorable than  5S1 $\rightarrow$ D despite the state lifetime being substantially shorter than the AT-all equivalent.

@@

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=450, scale=1]{Fig8.pdf}
        \caption{Diffusion map representation configurations sampled from the F4 state. The x and y axes represent the the first two non-trivial diffusion map eigenfunctions. Color maps show the probability that conformations are clustered into the frayed macrostate, maximum distance between 4th AT basepairs, and the overlap between adjacent A:T basepairs and the GC core.}
        \label{fig:GC-core_dmaps}
	\end{center}
\end{figure}

\subsection{Structural diversity in the F4 state informs fraying dynamics} 
(might just want to move this to the SI)

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=300, scale=1]{Fig9.pdf}
        \caption{Sequence-dependent population distributions for trimmed hybridization events. $Psi_2$ represents the maximum structural variance across all sequence configurations and correlates strongly with hybridization progression. Populations are discretized in 200 bins and smoothed by averaging with the nearest 5 bins. Heat maps show the average intermolecular base distances in nm at increasing values of $Psi_2$, where the main diagonal shows native base pair distances.}
        \label{fig:Fig9}
	\end{center}
\end{figure}

\subsection{Differential fraying behavior determines direct hybridization pathways}

(I just completed most of this analysis on Friday. I think this shows a more direct comparison to the Tokmakoff kinetic model -- mainly all-or-nothing deviations caused by fraying are ordered GC-core > GC-mix > AT-all > GC-end. In addition we see a distinct state form for GC-core which serves to verify the F4 state and its kinetic relevance. Lastly it provides another means by which MSMs can inform analysis techniques for clearer comparison between sequences. I could repeat this procedure and analysis for direct dissociation as well if that is of interest)

To more directly compare hybridization events across all sequences, we discarded transitions that passed through long-lived out-of-register states. To achieve higher resolutions, we reran simulations centered on direct D$\rightarrow$H (or D$\rightarrow$F4$\rightarrow$D for GC-core) transitions with a 10x higher save rate. We built diffusion maps on a set of 48 such "trimmed" events evenly sampled across sequences where each event consisted of 1000 frames. We found a distinct spectral gap after the first non-trivial diffusion map eigenvalue, indicating that the leading mode $\Psi_2$ contained most of the structural variance. Indeed, we found $\Psi_2$ had a 0.89 Pearson correlation with simulation time a -0.94 correlation with strand core distance with minimal variation by sequence. This shows that $\Psi_2$ is strongly correlated with the hybridization process for all sequences and represents a consistent collective variable for comparison between sequences. For each sequence, we compared population distributions along $\Psi_2$ and noted statistically meaningful differences in the range $ 0.001 <\Psi_2 < 0.0035$ (Figure \ref{fig:GC-core_dmaps}). In this region, some base pair contacts have formed but frayed structures dominate the distribution. As expected, GC-core shows a distinct peak in the frayed region, corresponding to the F4 macrostate identified by SRV-MSMs. GC-mix has a notably higher population than AT-all and GC-end but displays a more gradual progression toward the native hybridized state. This is consistent with a the downhill "nucleation-zippering" mechanism noted above.

@@

In any high-level model there are inevitable simulation artefacts produced by coarse-grained approximations. For example, the treatment of non-interacting base pairs as an excluded volume potential alone may not be representative of dynamics produced from mismatched dangling ends. In general, coarse-grained models produce a smoother free-energy surface which can results in must faster motions between states. This is illustrated by substantial accelerations in temperature-dependent responses when compared to experiment. Furthermore, it may be easier to cross between states -- e.g. a 5S1 $\rightarrow$ H transition -- when the usually rough free energy path becomes more easily traversed.

@@

\subsection{SRV training}

@@

\subsection{SRV-MSM walkthrough and analysis}
In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the most discernable spectral gap between modes. For this reason, we use this sequence a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that most implied timescales converge at an SRV lag time of 1.2 ns. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in figure \ref{fig:allseq_srv_crossval}), we see diminishing returns in the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis.

@@

>>>>

% need to discuss tmelt shift
We measured the slow dissociation response by fitting the distribution of times at which the core base pairs separate beyond a 2 nm cutoff. We found the inverse of these relaxation times -- the effective dissociation rate -- to increase exponentially with temperature, which is expected given the large enthalpic barrier of dissociation \citep{Craig1971ElaxationOligon,Porschke1971CooperativeTransition,Williams1989LaserDGCATGC}. We noted an average temperature shift of about 4C between experiment and simulation exponentials. While 3spn2 captures melting temperature relatively well, some deviation is expected given varied ion conditions and other coarse-grained effects. We used to this temperature shift to inform our equilibrium simulations in the proceeding section. We also noted an acceleration of about one order of magnitude compared to experiment, although this factor is sensitive to the exact definition of melting temperature. After accounting for the melting temperature shift and acceleration factor, we saw strong agreement between the experimental data and simulated relaxation fits. At lower temperatures, T-jump slow responses likely contain a mix of dissociation and hybridization dynamics, therefore rates do not drop off to the same extent as our analysis which considers dissociation alone. GC-core showed the largest deviation from experiment, indicating that increased temperature triggers dissociation at a disproportionately higher rate in simulation compared to experiment.

@@

For A:T terminal sequences, the simulated fast response appears linear with temperature, indicating a barrierless and diffusion-driven process. In contrast, the GC-end responses are distinctly slower and increase exponentially with temperature, likely due to a greater enthalpic barrier associated with G:C fraying. We observe similar trends in the experimental data, although comparisons at high temperatures were limited due to mixing between the fast and slow responses (figure \ref{fig:relaxation-comparison}). We found the optimal acceleration factor to be dependent on the choice of the cutoff parameter, but we can approximate that 3spn2 fraying dynamics are accelerated by about 100-200x relative to experiments. It is not surprising that we see different rates of acceleration for the dissociation and fraying processes given that coarse-grained effects can vary across different degrees of freedom. In particular, the simplified treatment of the fraying process may smoothen the free energy landscape and speed up dynamics relative to a more global processes like dissociation. Although one should rely on higher resolution models to study in depth mechanisms of fraying, our results indicate that terminal fraying is a reasonable assignment for the spectroscopic fast response in Sanstead et al. and that 3spn2 can capture some sequence-dependent fraying effects.
% need a better transition here back into SRV-MSMs?

@@

For GC-end we observe no significant flux between the 5S2 state and the structurally similar 5S1 state, indicating that the former acts more as a kinetic trap than a pathway to H. The 5S1 and 3S1 states still readily convert with H, however there is an order of magnitude lower probability of reaching these states from D when compared to AT-all. This suggests that inert tails inhibit out-of-register binding -- accounting in part for higher free energies discussed above -- but may not substantially disrupt out-of-register hybridization mechanisms along 5S1 $\rightarrow$ H and 3S1 $\rightarrow$ H once shifted states have formed. This phenomenon will be explored more in depth below.

% This mechanism is explored more in depth later on?

@@

Taken together, our results show that AT-all hybridization and dissociation kinetics are substantially modulated by out-of-register pathways. A majority of initial hybridization contacts occur out-of-register, and, based on transition probabilities, 35\% of D $\rightarrow$ H pathways and 37\% of H $\rightarrow$ D pathways pass through some out-of-register state. The 3S2 and 5s2 states are more prone to re-dissociation and thus contribute to about 5\% of these pathways. While GC-end strands can access out-of-register states, only 11\% of native hybridization events pass through any out-of-register states. These observations show that small disruptions to repetitive tracts shift hybridization pathways towards more two-state behavior. Indeed, there remain some repetitive tracts in GC-mix and GC-core, however we observe no out-of-register pathways for these sequences. Furthermore, our results indicate that internal displacement mechanisms reported in previous simulation studies are capable of disrupting non-native pairing and correcting base pair mismatches without fully dissociating strands \citep{Romano2013DNADependence, Markegard2015, Maciejczyk2014DNAModel}. These processes can occur even when terminal mismatches are present in the non-native structure.

Experimentally, out-of-register states are suspected to occur, but are difficult to capture due to short lifetimes and similarity in response to other processes such as fraying. To explicitly minimize out-of-register base pairing, similar AT repeat motif sequences have been padded by GCG clamps during experimental analysis \cite{Wyer2014KineticsAT-tracts}. Recent all-atom results identified analogous out-of-register states as "deep kinetic traps" along the hybridization pathway for repetitive dGCGCGC hexamers \citep{Xiao2019}. Contrary to our kinetic results, however, "slithering" mechanisms along out-of-register tracts were not observed to be a dominant pathway, especially when compared to the high rates of slithering exhibited by the homogenous dGGGGGG strand. It is unclear whether these differences were a consequence of varied simulation conditions or whether GC repeat motifs are less susceptible to direct out-obref-register transitions compared to AT motifs. This is possible given that stronger hydrogen binding in GC motifs \citep{Yakovchuk2006Base-stackingHelix,Zacharias2020Base-PairingFormation} may prevent fluctuation-driven rearrangement, however further computational and experimental studies are required to verify these differences.

\subsection{Out-of-register dynamics may stretch T-jump IR responses}

@@

% discuss computational comparisons needs work
Lattice model studies have shown that frayed intermediates make substantial contributions to the GC-core conformational ensemble \citep{Araque2016LatticeCooperativity, Phys2019,Sanstead2020OxidizedDNA}. Araque et al. defines a similar 8-mer sequence (dATGCGCAT) as non-two-state, where a stable, symmetrically A:T frayed state is a crucial part of the duplex transition path \citep{Araque2016LatticeCooperativity}. When examining all four sequences using T-jump IR and 2D IR spectroscopy, Sanstead et al. found that the GC-core had the highest deviation from two-state behavior during dissociation (when excluding out-of-register contributions) \citep{Sanstead2016}. As T-jump IR data only showed a loss of A:T contacts during the fast response, this intermediate state was defined by a high degree of fraying about the central core. While 1-2 base pair fraying was commonly observed for GC-mix and AT-all as well, lattice model predictions showed that GC-core had substantially more frayed base pairs \citep{Phys2019}. Variable T-jump measurements and Smoluchowski simulations on model 1D free energy landscapes showed that AT termini fraying was an effectively barrierless process characterized by rapid inter-conversion between all accessible frayed states \citep{Sanstead2018DirectDehybridization}. We see the same rapid fraying in simulation data -- which is too fast to be attributed to a converged SRV mode -- however our SRV-MSM results show this inter-conversion first relies on the slower formation of the the A:T bond nearest to the GC center.  Although this process occurs much slower than single A:T base bonding and breaking, it may be difficult to experimentally discern from the overall hybridization process which contains both G:C and A:T character and occurs on a similar timescale.

@@

@@

Although our SRV-MSMs provide useful insight into slow processes, we were also interested in the ensemble of configurations within PCCA+ macrostates and during transition periods. These configurations interchange significantly faster than the SRV and MSM lag times, but we can use a diffusion map embedding to glean a structural understanding of the macrostate population. Diffusion maps generate a low dimensional embedding of the data into high variance structural modes and are well-suited to find subtle differences in temporally disconnected data \citep{Coifman2006DiffusionMaps, Ferguson2010SystematicMaps}.

@@

The first diffusion mode clearly delineates between the GC-end and AT-all shifted conformations and correlates highly with the average distance between the 3' end and its shifted complementary pair. As expected, this shows that mismatched C:T pairs are never bound whereas the AT-all pairs are mostly bound with occasional fraying indicated by small AT-all overlap in the GC-end region. The second diffusion mode, which correlates highly with the average distance between 3' and 5' ends, highlights structural variances within the two sequences distributions. We notice that about 40\% of GC-end configurations fall below $\Psi_3 = -0.01$  compared to only 10\% of AT-all configuration. We find that these GC-end populations share stabilizing termini contacts despite all A:T contacts being shifting out-of-register. These "shifted-loop" bonds appear uniquely stable for GC-end conformations,  and suggest an intermediate states through which hybridization is facilitated. This may account for the GC-end 5S1 $\rightarrow$ H transition being more favorable than  5S1 $\rightarrow$ D despite the state free energy being substantially higher than the AT-all equivalent.

@@

\subsection{Sequence-dependent fraying determines direct hybridization pathways}

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=300, scale=1]{Fig8.pdf}
        \caption{Sequence-dependent population distributions for trimmed hybridization events. $Psi_2$ represents the maximum structural variance across all sequence configurations and correlates strongly with hybridization progression. Populations are discretized in 200 bins and smoothed by averaging with the nearest 5 bins. Heat maps show the average intermolecular base distances in nm at increasing values of $Psi_2$, where the main diagonal shows native base pair distances.}
        \label{fig:Fig8}
	\end{center}
\end{figure}

(I just completed most of this analysis on Friday. I think this shows a more direct comparison to Paul's 2016 kinetic model -- mainly all-or-nothing deviations caused by fraying are ordered GC-core > GC-mix > AT-all > GC-end. In addition we see a distinct state form for GC-core which serves to verify the F4 state and its kinetic relevance. Lastly it provides another means by which MSMs can inform analysis techniques for clearer comparison between sequences. I could repeat this procedure and analysis for direct dissociation as well if that is of interest)

To more directly compare fast hybridization dynamics across all sequences, we trimmed simulations centered on direct D$\rightarrow$H (or D$\rightarrow$F4$\rightarrow$D for GC-core) events and excluded transitions that passed through long-lived out-of-register states. To achieve higher resolutions, we reran simulations with a 10x higher save rate and selected a 1000 frame region (10 ns simulation time) for each event. We built diffusion maps on a set of 48 such "trimmed" events evenly sampled across sequences. We found a distinct spectral gap after the first non-trivial diffusion map eigenvalue, indicating that the leading mode $\Psi_2$ contained most of the structural variance. Indeed, we found $\Psi_2$ had a 0.89 Pearson correlation with simulation time and a  -0.94 correlation with inter-strand core distance with minimal variation by sequence. This shows that $\Psi_2$ is strongly correlated with hybridization progress for all sequences and represents a consistent collective variable for comparison between sequences.

For each sequence, we compared population distributions along $\Psi_2$ and noted statistically meaningful differences in the range $ 0.001 <\Psi_2 < 0.0035$ (Figure \ref{fig:Fig8}). In this region, some base pair contacts have formed but frayed structures dominate the distribution. As expected, GC-core shows a distinct peak in the frayed region, corresponding to the F4 macrostate identified by SRV-MSMs. GC-mix has a notably higher population than AT-all and GC-end but displays a more gradual progression toward the native hybridized state. This is consistent with the downhill "nucleation-zippering" mechanism described above. Although differences between AT-all and GC-end are more subtle, overall sequence populations in this region follow predictions by Sanstead et. al. regarding fdeviations from two-state binding behavior, where GC-core > GC-mix > AT-all > GC-end \citep{Sanstead2016}. This qualitative agreement shows that our MSM information can be adapted to identify dynamic behavior within the lag time and explore different definitions of two-state deviations.

@@

In any high-level model there are inevitable simulation artefacts produced by coarse-grained approximations. For example, the treatment of non-interacting base pairs as an excluded volume potential alone may not be representative of dynamics produced from mismatched dangling ends. In general, coarse-grained models produce a smoother free-energy surface which can results in much faster motions between states. This is illustrated by substantial accelerations in temperature-dependent responses when compared to experiment. Furthermore, it may be easier to cross between states -- e.g. a 5S1 $\rightarrow$ H transition -- when the usually rough free energy path becomes more easily traversed.

@@

\subsection{SRV hyperparameters}

@@

\subsection{SRV-MSM walkthrough and analysis}
In our analysis, we found that the AT-all sequence, given its repetitive structure and lack of GC-content, produced the most discernable spectral gap between modes. For this reason, we use this sequence a case study to work through our SRV-MSM pipeline step-by-step. Our first task was to identify the SRV lag time that was longer than the intrinsic Markov timescales of the system, yet short enough to resolve the dynamics of interest \citep{Phys2018MarkovValidation}. We found that an SRV lag time of 1.2 ns showed good convergence across the relevant timescales for all sequences. Next we selected an optimal number of SRV components to include in our analysis. After a certain point, higher order dynamical modes provide diminishing contributions the overall kinetic variance as measured by the Vamp-2 score, and the model can begin fitting on statistical noise in the trajectory data instead of the true dynamics \citep{McGibbon2015VariationalKinetics}. It is also more difficult to perform kmeans clustering on a high dimensional space, especially when those higher dimensions are less kinetically relevant \citep{Pande2010EverythingAsk}. For these reasons, the number of slow SRV components should be carefully selected based on the specific system of interest. As shown in figure \ref{fig:allseq_srv_crossval}), we see diminishing returns in the VAMP-2 score after five slow modes and select these modes as our optimized SRV basis.

@@

\subsection{Structural diversity in the F4 state informs fraying dynamics}

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=450, scale=1]{Fig9.pdf}
        \caption{Diffusion map representation configurations sampled from the F4 state. The x and y axes represent the the first two non-trivial diffusion map eigenfunctions. Color maps show the probability that conformations are clustered into the frayed macrostate, maximum distance between 4th AT basepairs, and the overlap between adjacent A:T basepairs and the GC core.}
        \label{fig:GC-core_dmaps}
	\end{center}
\end{figure}
